%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

% \documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
% \usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
% \usepackage{mathptmx} % assumes new font selection scheme installed

\let\proof\relax
\let\endproof\relax

\usepackage{amsthm}
\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage[usenames, dvipsnames, table]{xcolor}
\usepackage{float}
\usepackage[style=ieee,sorting=none,backend=biber, maxbibnames=99]{biblatex}
\addbibresource{bib.bib}
\usepackage[hidelinks]{hyperref}
\usepackage[ruled,resetcount,vlined]{algorithm2e}
\usepackage{commath}
\usepackage[font=small]{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{diagbox}
% \usepackage{transparent}
% \usepackage{xcolor}
\definecolor{dusty_blue}{HTML}{56B4E9}
\definecolor{dusty_green}{HTML}{009E73}
\definecolor{dusty_pink}{HTML}{CC79A7}

\colorlet{mylinkcolor}{BrickRed}
\colorlet{mycitecolor}{Green}
\colorlet{myurlcolor}{NavyBlue}

\hypersetup{
  linkcolor  = mylinkcolor,
  citecolor  = mycitecolor,
  urlcolor   = myurlcolor,
  colorlinks = true,
}

\newcommand{\ar}[1]{{\color{ForestGreen}AR:#1} }
\newcommand{\ap}[1]{{\color{purple}AP:#1} }
\newcommand{\ava}[1]{{\color{blue}ava:#1} }
\newcommand{\gm}[1]{{\color{red}GM:#1} }

\usepackage[capitalize]{cleveref}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\renewcommand{\Cref}[1]{\cref{#1}}
\Crefname{definition}{Def.}{Defs.}

\title{\LARGE \bf
Exploring How Non-Prehensile Manipulation Expands Capability in Robots Experiencing Multi-Joint Failure
}


\author{Gilberto Briscoe-Martinez$^{*}$, Anuj Pasricha, Ava Abderezaei, Santosh Chaganti,\\ Sarath Chandra Vajrala, Sri Kanth Popuri, and Alessandro Roncone% <-this % stops a space
\thanks{$^{*}$ Corresponding author.}
\thanks{GBM is supported by NASA Space Technology Graduate Research
Opportunity Grant 80NSSC22K1211.}% <-this % stops a space
\thanks{All authors are with the Department of Computer Science, University of Colorado Boulder, 1111 Engineering Drive, Boulder, CO USA.
{\tt\small firstname.lastname@colorado.edu}}
}



\begin{document}



\maketitle
\thispagestyle{empty}
\pagestyle{empty}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
This work explores non-prehensile manipulation (NPM) and whole-body interaction as strategies for enabling robotic manipulators to conduct manipulation tasks despite experiencing locked multi-joint (LMJ) failures. LMJs are critical system faults where two or more joints become inoperable; they impose constraints on the robot's configuration and control spaces, consequently limiting the capability and reach of a prehensile-only approach. This approach involves three components: i) modeling the failure-constrained workspace of the robot, ii) generating a kinodynamic map of NPM actions within this workspace, and iii) a manipulation action planner that uses a sim-in-the-loop approach to select the best actions to take from the kinodynamic map. The experimental evaluation shows that our approach can increase the failure-constrained reachable area in LMJ cases by 79\%. Further, it demonstrates the ability to complete real-world manipulation with up to 88.9\% success when the end-effector is unusable and up to 100\% success when it is usable.
\end{abstract}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \input{intro}
\input{new_intro}
% \input{anuj_intro}
\input{new_related}
% \input{adaptr_related}
\input{methods}
\input{Evaluation}
\input{conclusion}

\AtNextBibliography{\small}
\printbibliography

\end{document}
