\subsection{Multi-Query Planning via Lazy Edge Bundles}
\label{sec:discussion:multiquery}

\cref{ch:multiquery} addressed a practical limitation of simulation-based kinodynamic planning: dynamic feasibility is expensive to evaluate repeatedly.
In single-query planners, each new problem typically requires propagating candidate controls, checking collisions, and constructing feasible motion structure from scratch.
This is acceptable for isolated demonstrations, but limits deployment in settings where a robot must solve many related planning problems in the same workspace.

\textit{LazyBoE} shifts part of this computation offline by precomputing control-parameterized edge bundles over discretized state and control spaces.
These bundles encode reusable local motion structure, while lazy propagation and collision checking defer expensive evaluations until they are needed for a candidate solution.
The resulting planner preserves the flexibility of kinodynamic search while amortizing repeated dynamics computation across queries.

The broader contribution is a change in how dynamic feasibility is represented.
Rather than treating every planning query as an independent search through continuous dynamics, \textit{LazyBoE} stores reusable structure that can be selectively activated at planning time.
This makes multi-query kinodynamic planning more practical and establishes an intermediate point in the thesis arc: dynamics are no longer only evaluated online through search, but are partially represented, reused, and amortized across related problems.

% \subsection{Multi-Query Planning via Lazy Edge Bundles}
% \label{sec:discussion:multiquery}

% A practical bottleneck for deploying kinodynamic planners is that repeated queries over a shared workspace each incur the full cost of dynamics propagation and collision checking.
% \textit{LazyBoE} (\cref{ch:multiquery}) addresses this by pre-computing a graph of control-parameterized edge bundles over discretized state and control spaces, then applying lazy evaluation: deferring propagation and collision checking until an edge is actually required for a solution.
% Across multi-query manipulation benchmarks, \textit{LazyBoE} reduces average planning time and achieves higher success rates than existing baselines by exploring a wider range of solutions without paying the full propagation cost upfront.
% The broader takeaway is that for kinodynamic planning to be practically deployable---rather than a one-off demonstration---amortizing cost across queries is at least as important as reducing per-query complexity.