\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amsmath}		% math environments and \text{}
\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??
\usepackage[ruled,linesnumbered,resetcount,vlined]{algorithm2e}	% algorithm pseudocode
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{multirow}		% multi-row table cells
\usepackage{bm}			% bold math via \bm{}
\usepackage{wrapfig}		% text-wrapped figures
\usepackage[usenames,dvipsnames]{xcolor}	% named colors and \definecolor{}{HTML}{}
\usepackage{soul}			% \ul{} underline for emphasis

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Expanding Robot Manipulation Capability Through Dynamics-Constrained Motion Generation}

% TODO: Personalize the front matter below.
\author{Anuj}{Pasricha}

\otherdegrees{M.S., University of Colorado Boulder, 2019 \\B.S., University of Illinois at Urbana-Champaign, 2015}

\degree{Doctor of Philosophy}		%  #1 {long descr.}
	{Ph.D., Computer Science}		%  #2 {short descr.}

\dept{Department of}			%  #1 {designation}
	{Computer Science}		%  #2 {name}

\advisor{Prof.}				%  #1 {title}
	{Alessandro Roncone}			%  #2 {name}

\reader{Bradley Hayes}		%  2nd person to sign thesis
\readerThree{Nikolaus Correll}	%  3rd person to sign thesis
\readerFour{Jeffrey Ichnowski}
\readerFive{Mehmet Dogar}

\abstract{  \OnePageChapter	% remove \OnePageChapter once the abstract grows past a page.

Humans carry implicit knowledge of what motions are physically permissible: how to slide a heavy object without tipping it, how to carry a liquid-filled container without spilling, how to adjust effort when a load is unexpectedly heavy.
Robotic systems must represent that knowledge explicitly---and the dominant paradigm, which organizes manipulation around geometric reasoning, grasp synthesis, and collision avoidance, makes that representation implicit in the algorithm itself rather than treating it as a first-class design concern.
The result is a class of manipulation behaviors---impulsive non-prehensile rearrangement, high-payload handling, liquid transport, operation under actuator failure---that geometric abstractions cannot reliably plan for, because the constraints that determine feasibility are dynamic, not kinematic.

This thesis argues that knowledge of dynamic feasibility---what motions are physically possible given a robot's constraints and the task physics---should be treated as a design object.
Where that knowledge resides in a robotic system, what computational form it takes, and at what point in the planning process it is evaluated are primary decisions that determine what manipulation behaviors become possible.
By presenting three different answers to these questions, this thesis makes it possible to design manipulation systems in which feasibility is not an implicit consequence of the chosen algorithm, but an explicit object of representation, reuse, and learning.

The first part places constraint knowledge online, in the forward model of a physics-simulation-based kinodynamic planner, enabling non-prehensile rearrangement via impulsive contact, whole-body manipulation under joint failures, and spill-free liquid transport.
The second part amortizes that knowledge into reusable precomputed motion structures, reducing repeated planning cost while preserving physical grounding.
The third part absorbs constraint knowledge into learned conditional trajectory distributions---diffusion-based generators conditioned on payload mass and embodiment configuration---and extends this representation to near-feasible seed generation for constrained trajectory optimization.
Together, these contributions demonstrate that model-based and learning-based approaches to manipulation are not competing paradigms but different mechanisms for placing constraint knowledge, and that choosing among them is itself a consequential design decision.
	}

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	To the one stranger who might read this think new thoughts.
	}

\acknowledgements{	\OnePageChapter	% *MUST* BE ONLY ONE PAGE!
	Confidential.
	}

% \IRBprotocol{E927F29.001X}	% optional!

\ToCisShort	% use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input macros.tex
% \input chapter1.tex
\input{ThesisFilePkg/ch1}
% \input chapter2.tex
\input{ThesisFilePkg/ch1mid2}
\input chapter3.tex
\input chapter4.tex
\input chapter5.tex
\input chapter6.tex
\input chapter7.tex
\input chapter9.tex
\input chapter10.tex

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{plain}	% or "siam", or "alpha", etc.
\nocite{*}		% list all refs in database, cited or not
\bibliography{refs}		% Bib database in "refs.bib"

%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
% \appendix
% \input appendixA.tex


\end{document}

