\section{Payload-Conditioned Trajectory Denoising}

In this section, we present our payload-conditioned diffusion model to generate pick-and-place trajectories that transport super-nominal payloads while obeying kinematic and dynamic constraints.

\subsection{Training Data}

Our payload-conditioned diffusion model (\Cref{sec:model}) is trained on $25,000$ time-parameterized, collision-free joint space trajectories (state dimension = DoF$\times3$, corresponding to joint angles, velocities, and accelerations) for tabletop pick-and-place motions. These trajectories are generated using a plan-and-filter pipeline that uses cuRobo, a GPU-accelerated trajectory optimization framework \cite{sundaralingam2023curobo}, followed by a dynamics validation step (\Cref{fig:data-generation}).
\input{figs/data-generation}
First, trajectories are generated for the 7DoF Franka Emika Panda robot arm with an effective payload mass of $0$kg.
Second, to ensure that payload-induced joint torques remain within hardware limits, each trajectory is verified through an analytically-derived dynamics model parameterized on real-world data \cite{gaz2019dynamic}.
This model captures both the standard dynamic elements (inertia, Coriolis, and gravity) and velocity-dependent joint friction.
By using a dynamics model fitted to real robot trajectory data, we ensure that our training dataset reflects the physical constraints observed in hardware.

This dynamics verification step utilizes a complete robot model where $q$, $\dot{q}$, $\ddot{q}$, and $\tau$ are $7\times1$ vectors representing joint angles, velocities, accelerations, and torques respectively.
The dynamics matrices include: $M(q)$ ($7\times7$ positive-definite, symmetric joint inertia matrix), $C(q,\dot{q})$ ($7\times7$ Coriolis/centrifugal matrix), $g(q)$ ($7\times1$ gravity compensation torque vector), and $f(\dot{q})$ ($7\times1$ joint friction torque vector). External task space forces $F_{ext}$ ($6\times1$) map to joint torques through the Jacobian $J(q)$ ($7\times7$).
The dynamics model is parameterized as:
\begin{equation}\label{eq:arm-dyns}
    \tau = M(q)\ddot{q} + C(q, \dot{q})\dot{q} + g(q) + f(\dot{q}) + J^{-1}(q)F_{ext}
\end{equation}
Note that this analytically-derived model represents the robot without external payloads which can affect model accuracy when considering varying payload masses in real-world deployment.
To account for payload-induced joint torques, we model the gravitational effects as a 6D Cartesian wrench transformed from the world frame to the end-effector frame.
This frame transformation is critical for accurately projecting gravitational forces acting on the payload along the robot's embodiment.
We represent the external gravitational wrench $\bm{F}_g \in \mathbb{R}^6$ in the world frame with a simplified point mass model:
$\bm{F}_g = mg \cdot [0,\ 0,\ -1,\ 0,\ 0,\ 0]^T$
where $m \in \mathbb{R}^+$ denotes payload mass and $g \approx 9.81 \text{ m/s}^2$ is gravitational acceleration. This representation assumes force application at the object's center of mass, neglecting potential moment contributions.
The wrench's structure combines linear forces (first three components) and angular forces (last three components), which are
mapped to induced joint torques through the manipulator Jacobian, contributing to the total joint torque vector (\cref{eq:arm-dyns}).

The training dataset $\mathcal{D}$ is constructed from these trajectories that satisfy both kinematic and dynamic constraints. Each element $d_i \in \mathcal{D}$ is a tuple of trajectory $\bm{\pi}_i$ and maximum supported payload $m_i$. %, where $\bm{\pi}_i$ is as defined in \Cref{sec:diffusion}.
Each $m_i$ is the maximum supported payload mass for trajectory $i$ subject to $|\tau(\bm{\pi}_i, m_i)| \leq \tau_{max}$, $\forall t \in [0,T]$ where $\tau_t$ is computed via the dynamics equation (\cref{eq:arm-dyns}).
\input{3b.payload-conditioning}